---
title: "Tables Assignment"
author: "William Roloff"
date: "2022-10-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For this assignment I used a data set consisting of NBA player data, which I also used on the first two assignments.  I am presenting information specifically on 3-point shooting and how those numbers have changed over the years.  This table is filtered to include only players that took greater than 150 3-point shots in a season and made at least 40% of those shots.  The 3-point shot was introduced in 1980, and one of my goals with this table is to show how rare it was for a player to shoot a high volume and percentage of 3-pointers in the 80s compared to how commonplace it is in recent seasons.  The table is quite long, so I included fixed headers, and I enabled highlighting of rows to ensure that it's easier to keep track of the data as the user is scrolling.

```{r}
nba_stats <- read.csv("nba stats csv.csv")



library(tidyverse)
library(knitr)
#first5 <- head((nba_stats), n=5)
#kable(first5)

library(kableExtra)
#kbl(first5)
```


```{r}
#tbl_first5 <- first5 %>% 
#  select(Year, Player, X3P, X3PA, X3P., USG., WS) %>%
#  kbl(col.names = c("Year",
#                    "Player",
#                    "3-point Makes",
#                    "3-point Attempts",
#                    "3-point Percentage",
#                    "Usage Rate",
#                    "Win Shares"))  %>%
#  kable_styling(font_size=18)

#tbl_first5
```


```{r}
nba_data_3 <- subset(nba_stats, Year>1979)
nba_data_3_attempts <- subset(nba_data_3, X3PA>150)
nba_data_3_top <- subset(nba_data_3_attempts, X3P.>.4)
#head(nba_data_3_top)
```


```{r}
averages <- nba_data_3 %>%
  group_by(Year) %>%
  summarise_each(funs(mean))

#averages
```


```{r}
averages_simple <- averages %>%
                      mutate('X3P.' = X3P/X3PA)

#averages_simple

three_averages <- subset(averages_simple, select = c('Year', 'X3PA', 'X3P.'))
three_averages_names <- rename(three_averages, 'X3PA.mean' = X3PA, 'X3P.mean' = X3P.)
#head(three_averages_names)
```
```{r}
nba_data_3_w_avg <- right_join(three_averages_names, nba_data_3_top)
#head(nba_data_3_w_avg)
```


```{r}
tbl_3data <- nba_data_3_w_avg %>% 
  select(Year, Player, X3P, X3PA, X3PA.mean, X3P., X3P.mean) %>%
  kbl(col.names = c("Year",
                    "Player",
                    "3-point Makes",
                    "3-point Attempts",
                    "Average Attempts for that Year",
                    "3-point Percentage",
                    "Average 3-point Percentage for that Year"))  %>%
  kable_styling(font_size=18)

#tbl_3data
```

```{r}
tbl_3data %>%
  kable_paper(lightable_options = "hover", full_width = FALSE) %>%
  kable_styling(fixed_thead = TRUE)
```



